### 2011-04-07 docs/enthalpy/enthalpy.tex
was missing a factor of rho
fixed sign error and added some commentary

The mismatch between the pressure implicit in the definition of $h$
and $p_0$ can be seen by substituting $h = e + p/\rho$ into the
enthalpy equation, where we replace $p$ with $p_0$ in the $Dp/Dt$ term:
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + 
  \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} + \frac{\partial p}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) -\nabla\cdot(p\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) - p\nabla\cdot\Ub + \rho\Hnuc + 
  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} \nonumber 
\end{eqnarray}
However, if we solve the evolution equation for $e$ we would substitute 
$p_0$ for $p$ in this equation as well.  Thus, we can pose the situation as the following.  
If we solve the evolution equation for $h$ then we effectively are solving
\begin{equation}
\frac{\partial(\rho e)}{\partial t} +
  \nabla\cdot(\rho e\Ub) = p \; \nabla\cdot\Ub + \rho\Hnuc + 
  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} \nonumber 
\end{equation}
but if we solve the evolution equation for $e$ we are effectively solving
\begin{equation}
\frac{\partial(\rho e)}{\partial t} +
  \nabla\cdot(\rho e\Ub) = p_0 \nabla\cdot\Ub + \rho\Hnuc \nonumber 
\end{equation}

The second equation subtracted from the first gives:
\begin{equation} \label{eq:difference between h and e equations}
  \frac{D (p_0 - p)}{Dt} - (p_0 - p)  \nabla \cdot \Ub
\end{equation}

Suppose we solve the current enthalpy equation, but when we call the EOS, we 
subtract $p_0$ from $\rho h$ and then call the EOS with $e$ instead of $h$.  This is equivalent
to:
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) +
  \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} + \frac{\partial p_0}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) -\nabla\cdot(p_0 \Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) - p_0 \nabla\cdot\Ub + \rho\Hnuc  \nonumber
\end{eqnarray}
which is identical to solving the energy equation.

\subsection{Constant $\gamma$ Gas}
Going back to the constant $\gamma$, ideal gas EOS, we can rewrite the
enthalpy equation as a pressure evolution equation
\begin{eqnarray}
  \frac{\partial\rho h}{\partial t} + \nabla\cdot\left(\rho h\Ub\right) &=& \frac{Dp_0}{Dt} + \rho H {} \nonumber\\
  \frac{\gamma}{\gamma-1}\frac{\partial p}{\partial t} + \frac{\gamma}{\gamma-1}\nabla\cdot\left(p\Ub\right) &=& \frac{Dp_0}{Dt} + \rho H {} \nonumber\\
  \frac{Dp}{Dt} &=& -p\nabla\cdot\Ub + \frac{\gamma}{\gamma-1}\left(\frac{Dp_0}{Dt}+\rho H\right) \label{eq:H:pressure evolution}.
\end{eqnarray}
Similarly, we can derive a pressure evolution equation from the energy equation
\begin{equation}
  \frac{Dp}{Dt} = -p\nabla\cdot\Ub - \left(\gamma-1\right)p_0\nabla\cdot\Ub + \rho H \label{eq:e:pressure evolution}
\end{equation}
Now, if we further make the assumption that $p_0$ is constant,
$Dp_0/Dt = 0$, then the divergence constraint for such a gas reduces
to
\begin{equation}
\nabla\cdot\Ub = \frac{\gamma-1}{\gamma p_0}H \label{eq:div constraint for constant gamma}.
\end{equation}
Plugging this back into either of \eqref{eq:H:pressure evolution} or \eqref{eq:e:pressure evolution} gives
\begin{equation}
  \frac{Dp}{Dt} = \frac{\gamma-1}{\gamma}\left(1-\frac{p}{p_0}\right)H. \label{eq:pressure evolution constant gamma}
\end{equation}
If $p_0$ is assumed constant and using \eqref{eq:div constraint for
  constant gamma}, the difference between the enthalpy equation and
the energy equation, \eqref{eq:difference between h and e equations},
can be rewritten as
\[
-\frac{Dp}{Dt} - \frac{\gamma}{\gamma-1}\left(1-\frac{p}{p_0}\right)H = 0,
\]
where the equality holds from \eqref{eq:pressure evolution constant gamma}.

>>>>

The mismatch between the pressure implicit in the definition of $h$
and $p_0$ can be seen by substituting $h = e + p/\rho$ into the
enthalpy equation, where we replace $p$ with $p_0$ in the $Dp/Dt$ term:
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + 
  \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} + \frac{\partial p}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) -\nabla\cdot(p\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) - p\nabla\cdot\Ub + \rho\Hnuc + 
  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} \nonumber 
\end{eqnarray}
However, if we solve the evolution equation for $e$ we would substitute 
$p_0$ for $p$ in this equation as well.  Thus, we can pose the situation as the following.  
If we solve the evolution equation for $h$ then we effectively are solving
\begin{equation}
\frac{\partial(\rho e)}{\partial t} +
  \nabla\cdot(\rho e\Ub) = -p \; \nabla\cdot\Ub + \rho\Hnuc + 
  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} \nonumber 
\end{equation}
but if we solve the evolution equation for $e$ we are effectively solving
\begin{equation}
\frac{\partial(\rho e)}{\partial t} +
  \nabla\cdot(\rho e\Ub) = p_0 \nabla\cdot\Ub + \rho\Hnuc \nonumber 
\end{equation}

The second equation subtracted from the first gives:
\begin{equation} \label{eq:difference between h and e equations}
  \frac{D (p_0 - p)}{Dt} - (p_0 - p)  \nabla \cdot \Ub = 0,
\end{equation}
but this equation is only true, in general, if $p=p_0$.

Suppose we solve the current enthalpy equation, but when we call the EOS, we 
subtract $p_0$ from $\rho h$ and then call the EOS with $e$ instead of $h$.  This is equivalent
to:
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) +
  \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} + \frac{\partial p_0}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) -\nabla\cdot(p_0 \Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) - p_0 \nabla\cdot\Ub + \rho\Hnuc  \nonumber
\end{eqnarray}
which is identical to solving the energy equation with $p\to p_0$.

\subsection{Constant $\gamma$ Gas}
Going back to the constant $\gamma$, ideal gas EOS, we can rewrite the
enthalpy equation as a pressure evolution equation
\begin{eqnarray}
  \frac{\partial\rho h}{\partial t} + \nabla\cdot\left(\rho h\Ub\right) &=& \frac{Dp_0}{Dt} + \rho H {} \nonumber\\
  \frac{\gamma}{\gamma-1}\frac{\partial p}{\partial t} + \frac{\gamma}{\gamma-1}\nabla\cdot\left(p\Ub\right) &=& \frac{Dp_0}{Dt} + \rho H {} \nonumber\\
  \frac{Dp}{Dt} &=& -p\nabla\cdot\Ub + \frac{\gamma}{\gamma-1}\left(\frac{Dp_0}{Dt}+\rho H\right) \label{eq:H:pressure evolution}.
\end{eqnarray}
Similarly, we can derive a pressure evolution equation from the energy equation
\begin{equation}
  \frac{Dp}{Dt} = -p\nabla\cdot\Ub - \left(\gamma-1\right)p_0\nabla\cdot\Ub + \rho H \label{eq:e:pressure evolution}
\end{equation}
Now, if we further make the assumption that $p_0$ is constant,
$Dp_0/Dt = 0$, then the divergence constraint for such a gas reduces
to
\begin{equation}
\nabla\cdot\Ub = \frac{\gamma-1}{\gamma p_0}\rho H \label{eq:div constraint for constant gamma}.
\end{equation}
Plugging this back into either of \eqref{eq:H:pressure evolution} or \eqref{eq:e:pressure evolution} gives
\begin{equation}
  \frac{Dp}{Dt} = \frac{\gamma-1}{\gamma}\left(1-\frac{p}{p_0}\right)\rho H. \label{eq:pressure evolution constant gamma}
\end{equation}
If $p_0$ is assumed constant and using \eqref{eq:div constraint for
  constant gamma}, the difference between the enthalpy equation and
the energy equation, \eqref{eq:difference between h and e equations},
can be rewritten as
\[
-\frac{Dp}{Dt} - \frac{\gamma}{\gamma-1}\left(1-\frac{p}{p_0}\right)\rho H = 0,
\]
where the equality holds from \eqref{eq:pressure evolution constant
  gamma}. In other words, for the constant $\gamma$ gas we have
$p=p_0$ as expected.
